Young's Experiment

### 2020-11-10 fdtd/em.tex
WIP 10Nov2020

  for k in range(ke - 1):
    hy[k] += 0.5 * (ex[k] - ex[k + 1])
\end{verbatim}
Finally, we plot the results.
\begin{verbatim}
plt.figure(figsize=(8, 8))
plt.subplot(211)
plt.plot(ex, color='k', linewidth=1)
plt.ylabel('E$_x$', fontsize='14')
plt.xticks(np.arange(0, ke + 1, step=20))
plt.xlim(0, 200)
plt.yticks(np.arange(-1, 1.2, step=1))
plt.ylim(-1.2, 1.2)
plt.text(100, 0.5, 'T = {}'.format(t), horizontalalignment='center')
plt.subplot(212)
plt.plot(hy, color='k', linewidth=1)
plt.ylabel('H$_y$', fontsize='14')
plt.xlabel('FDTD cells')
plt.xticks(np.arange(0, ke + 1, step=20))
plt.xlim(0, 200)
plt.yticks(np.arange(-1, 1.2, step=1))
plt.ylim(-1.2, 1.2)
plt.subplots_adjust(bottom=0.2, hspace=0.45)
plt.show()
\end{verbatim}
The results of the simulation are
\begin{figure}
\centering
\includegraphics[scale=0.5]{fdtd-1}
\caption{Propagation of a Gaussian pulse}\label{f2}
\end{figure}

\bibliographystyle{plain}
\bibliography{em}
\end{document}

>>>>

  for k in range(ke - 1):
    hy[k] += 0.5 * (ex[k] - ex[k + 1])
\end{verbatim}
Finally, we plot the results.
\begin{verbatim}
plt.figure(figsize=(8, 8))
plt.subplot(211)
plt.plot(ex, color='k', linewidth=1)
plt.ylabel('E$_x$', fontsize='14')
plt.xticks(np.arange(0, ke + 1, step=20))
plt.xlim(0, 200)
plt.yticks(np.arange(-1, 1.2, step=1))
plt.ylim(-1.2, 1.2)
plt.text(100, 0.5, 'T = {}'.format(t), horizontalalignment='center')
plt.subplot(212)
plt.plot(hy, color='k', linewidth=1)
plt.ylabel('H$_y$', fontsize='14')
plt.xlabel('FDTD cells')
plt.xticks(np.arange(0, ke + 1, step=20))
plt.xlim(0, 200)
plt.yticks(np.arange(-1, 1.2, step=1))
plt.ylim(-1.2, 1.2)
plt.subplots_adjust(bottom=0.2, hspace=0.45)
plt.show()
\end{verbatim}
The results of the simulation are shown in figure \ref{f2}
\begin{figure}
\centering
\includegraphics[scale=0.5]{fdtd-1}
\caption{Propagation of a Gaussian pulse}\label{f2}
\end{figure}

\section{An example of a 2d simulation}\label{s8}
We will extend the code in the previous section to two dimensions. Once again,
the code is a minor modification of the code in Houle in Sullivan's book
\cite{sullivan2013electromagnetic}.

We first begin by importing the necessary packages.
\begin{verbatim}
import numpy as np
import matplotlib.pyplot as plt
\end{verbatim}

Next we set up the physical parameters.
\begin{verbatim}
ddx = 0.01        # Cell size.
c = 3e8           # Speed of light in vacuum.
dt = ddx / 2*c    # Time step size.
epsz = 8.854e-12  # Permittivity of free space.
t0 = 20           # Mean magnitude of the pulse.
spread = 6        # Spread of the pulse.
\end{verbatim}

The simulation parameters are:
\begin{verbatim}
ie =  60
je =  60
ic = ie // 2
jc =  je // 2
ez =  np.zeros((ie,  je))
dz =  np.zeros((ie,  je))
hx =  np.zeros((ie,  je))
hy =  np.zeros((ie,  je))
gaz = np.ones((ie, je))
nsteps = 50

# Dictionary to track data at various times.
plotting_points = [
  {'label': 'a', 'num_steps': 20, 'data_to_plot': None},
  {'label': 'b', 'num_steps': 30, 'data_to_plot': None},
  {'label': 'c', 'num_steps': 40, 'data_to_plot': None},
  {'label': 'd', 'num_steps': 50, 'data_to_plot': None},
]
\end{verbatim}

The code to compute the electric and magnetic fields is:
\begin{verbatim}
def compute_fields(time_step):
  for j in range(1, je):
    for i in range(1, ie):
      dz[i, j] = dz[i, j] + 0.5 * (hy[i, j] - hy[i - 1, j] - hx[i, j] + hx[i, j - 1])

  dz[ic, jc] = np.exp(-0.5 * ((t0 - time_step) / spread) ** 2)

  for j in range(1, je):
    for i in range(1, ie):
      ez[i, j] = gaz[i, j] * dz[i, j]

  for j in range(je - 1):
    for i in range(ie - 1):
      hx[i, j] = hx[i, j] + 0.5 * (ez[i, j] - ez[i, j + 1])

  for j in range(je - 1):
    for i in range(ie - 1):
      hy[i, j] = hy[i, j] + 0.5 * (ez[i + 1, j] - ez[i, j])

  for plotting_point in plotting_points:
    if time_step == plotting_point['num_steps']:
      plotting_point['data_to_plot'] = np.copy(ez)
\end{verbatim}

We compute the fields for a certain number of time steps, defined in the
simulation parameters setup.
\begin{verbatim}
for t in range(1, nsteps + 1):
  compute_fields(t)
\end{verbatim}

Finally, we plot the results.
\begin{verbatim}
plt.rcParams['font.size'] = 12
plt.rcParams['grid.color'] = 'gray'
plt.rcParams['grid.linestyle'] = 'dotted'
fig = plt.figure(figsize=(8, 7))
X, Y = np.meshgrid(range(je), range(ie))

def plot_e_field(ax, data, timestep, label):
    """3d Plot of E field at a single time step"""
    ax.set_zlim(0, 1)
    ax.view_init(elev=20., azim=45)
    ax.plot_surface(X, Y, data[:, :], rstride=1, cstride=1, 
                    color='white', edgecolor='black', linewidth=.25)
    ax.zaxis.set_rotate_label(False)
    ax.set_zlabel(r' $E_{Z}$', rotation=90, labelpad=10, fontsize=14)
    ax.set_zticks([0, 0.5, 1])
    ax.set_xlabel('cm')
    ax.set_ylabel('cm')
    ax.set_xticks(np.arange(0, 61, step=20))
    ax.set_yticks(np.arange(0, 61, step=20))
    ax.text2D(0.6, 0.7, f"T = {timestep}", transform=ax.transAxes)

    ax.xaxis.pane.fill = ax.yaxis.pane.fill = ax.zaxis.pane.fill = False
    plt.gca().patch.set_facecolor('white')
    ax.text2D(-0.2, 0.8, "({})".format(label), transform=ax.transAxes)
    ax.dist = 11

for subplot_num, plotting_point in enumerate(plotting_points):
    ax = fig.add_subplot(2, 2, subplot_num + 1, projection='3d')
    plot_e_field(ax, plotting_point['data_to_plot'],
    plotting_point['num_steps'],
    plotting_point['label'])
    plt.subplots_adjust(bottom=0.05, left=0.10, hspace=0.05)

plt.show()
\end{verbatim}
The results of the simulation are shown in figure \ref{f3}.
\begin{figure}
\centering
\includegraphics[scale=0.5]{fdtd-2}
\caption{Propagation of a Gaussian pulse in 2d}\label{f3}
\end{figure}
\bibliographystyle{plain}
\bibliography{em}
\end{document}
